---
title: "DynamicModels"
author: "Steven Cognac"
date: '2022-05-03'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(deSolve)
library(here)
library(tidyverse)

source(here("R/dexppop.R"))
```

# additional parameters 

You can play a bit with changing your function to something that you can't integrate "by hand"

BUT we might want more parameters

to work with ODE,
parameters must all be input as a single list; similar to how we return multiple outputs from a function

see example below..lets add a carrying capacity

# R code with carrying capacity


```{r}
initialrabbits = 20

years = seq(from=1, to=100, by=2)

# run the solver
Ptime = ode(y = initialrabbits,
            times = years,
            func = dexppop,
            parms = c(0.01))

# not exactly a dataframe
head(Ptime)
```
```{r}
colnames(Ptime)=c("year","P")
```

```{r}
# this also means you need to extract just the data frame for plotting
ggplot(as.data.frame(Ptime),aes(year,P)) +
  geom_point() + 
  labs(y = "Population",
       "years")
```

```{r}
# this also works (of course function can be by order)
Ptime = ode(initialrabbits, years, dexppop,0.03)

colnames(Ptime) = c("year","P")

ggplot(as.data.frame(Ptime),aes(year,P)) + 
  geom_point() +
  labs(y = "Population",
       x = "years")
```


# Difference Equations 

What is ODE doing? (iterating in 'smart ways')

Similar to "difference equations"

## Difference Equations

Population models can be discrete (rather than continuous)
So we could implement them as difference equations and iterate


```{r iteratie}
source(here("R/discrete_logistic_pop.R"))

# notice how a for loop is used to iterate
# how many rabbits after 50 years given a growth of 0.1
# starting with 1 rabbit - but a carrying capcity of 500
discrete_logistic_pop
discrete_logistic_pop(P0=1, r=0.05, K=200, T=50)

# save results
discrete_result = discrete_logistic_pop(P0=1, r=0.05, K=200, T=50)

# lets also keep the parameters for use later
P0=1
r=0.05
K=200
T=50
```

# Differential Equation, Difference (Iteration by hand) comparison

Remember we have 3 ways now to calculate population

analytical solution - based on integration (exppop.R) BEST

using an ode solver for numerical approximation (exppop_play.R)

numerical integration using in discrete steps (discrete_logistic_pop.R)
```{r closed}
source(here("R/exppop.R"))

exppop(P0=P0, r=r, K=K, T=T)
analytic_result = exppop(P0=P0, r=r, K=K, T=T)
analytic_result
discrete_result

# why are they different
# look at trajectories
growth_result = data.frame(time=seq(from=1,to=100))

growth_result$Panalytic = growth_result$time %>% 
  map_dbl(~exppop( P0=1,
                   r=0.05,
                   K=200,
                   T=.x ))

growth_result$Pdiscrete = growth_result$time %>%
  map_dbl(~discrete_logistic_pop(P0=1,
                                 r=0.05,
                                 K=200,
                                 T=.x ))

tmp = growth_result %>% 
  gather(key="Ptype",value="P",-time)

ggplot(tmp, aes(time,P, col=Ptype)) +
  geom_point()
# try running them for longer time periods to see what happens 
# change the value of r, K , P0 - see how it effects the results
```

# Compare analytical, difference and ODE

Finally look at continuous derivative using ODE solve
Needs
*initial condtions
*differential equation
*parameters

```{r}
source(here("R/dexppop_play.R"))

# set up using the same parameters
pcompare = list(r=r,carry_capacity=K)

# now run our ODE solver
result = ode(P0, growth_result$time, dexppop_play, pcompare)
head(result)

# we already have time - so just extract population
growth_result$Pdifferential = result[,2]

# comapre all 3 approaches
tmp = growth_result %>% 
  gather(key="Ptype",value="P",-time)

ggplot(tmp, aes(time,P, col=Ptype)) +
  geom_point()
# notice Pdifferential is closer to Panalytic than Pdiscrete
```

```{r}
# create parameter list
initalrabbits=2
newparms = list(r=0.03, carry_capacity=300)

#apply solver
results = ode(initialrabbits, years, dexppop_play,newparms)
head(results)

```



