---
title: "carbon_storage"
author: "Steven Cognac"
date: '2022-05-12'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(deSolve)
library(tidyverse)
library(here)
source(here("R/dharvest.R"))
source(here("R/fixed_harvest.R"))
```


```{r}

```


```{r carbon_forest_stock}
# given some forest characteristics - lets look at derivatives under different harvest rates
lowHrate = 0.015
gps = list(harv=lowHrate, K=100, r=0.05)

# look at the derivative over a range of forest sizes

findstable = data.frame(Ccurr=seq(from=1, to=100, by=5))
# notice use of Time=NULL, and map to compute derivative
# for different values of forest biomass
findstable$dervHlow= unlist(findstable$Ccurr %>% map(~dharvest(parms=gps, Time=NULL, P=.x) ))
                                                  
ggplot(findstable, aes(Ccurr, dervHlow)) +
  geom_point() +
  geom_hline(yintercept = 0, col="red") +
  labs(y="Derivative\n (Rate of change of forest carbon) (kg/C/year)", 
       x="Current Forest Carbon Stock (kgC)")
```
Interpreting graph. If you have 50 units of carbon, your carbon stock is growing. If you have 100 units of carbon, stock is shrinking, likely due to harvesting. Values on X-axis are hypothentical initial conditions. Given an initial harvest rate, if we harvest xyz, are my carbon stocks going to be stable?

```{r}
# Harvest will be stable when derivative is zero!

# look at a different harvest rate
midHrate=0.02
gps = list(harv=midHrate, K=100, r=0.05)
findstable$dervHmid = unlist(findstable$Ccurr %>% 
                              map(~dharvest(parms=gps, Time=NULL, P=.x) ))
 
# try high rate
highHrate=0.05
gps = list(harv=highHrate, K=100, r=0.05)
findstable$dervHhigh = unlist(findstable$Ccurr %>% 
                                map(~dharvest(parms=gps, Time=NULL, P=.x) ))
 

# plot them all together
tmp = gather(findstable, key="HarvestRate", value="value", -Ccurr)

ggplot(tmp, aes(Ccurr, value, color=HarvestRate)) + 
  geom_point() +
  geom_hline(yintercept = 0, col="black") +
  labs(x="Forest Biomass (kgC)",
       y="Forest Growth Rate (kgC/year)")
```

If we start with a Forest Carbon of 50 and low harvest rate what will happen

If we start with a Forest Carbon of 50 and a high harvest rate what will happen


```{r}
tm = seq(from=1, to=500)
gps = list(harv=lowHrate, K=100, r=0.05)
Pinitial=50
res = ode(Pinitial,tm, dharvest, gps)
colnames(res)=c("time","carbon")

# plot
ggplot(as.data.frame(res), aes(time, carbon))+geom_point()+
  labs(y="Forest Biomass (kgC)", x="Year", title="low harvest rate")
```

```{r}
gps = list(harv=highHrate, K=100, r=0.05)
Pinitial=50
res = ode(Pinitial,tm, dharvest, gps)
colnames(res)=c("time","carbon")

# plot
ggplot(as.data.frame(res), aes(time, carbon))+geom_point()+
 labs(y="Forest Biomass (kgC)", x="Year", title="High Harvest Rate")

```


```{r fixed_rate}
#forest growth rate of 0.05, a carrying capacity of 1000kg and a harvest rate of 10kg/year
carbon = seq(from=0, to=1000)

metrics = list(r=0.05, K=1000, mincarbon=0, harv=10)

dcarbon= unlist(carbon %>% 
                  map(fixed_harvest, Time=NULL, parms=metrics))

ggplot(as.data.frame(carbon,dcarbon), aes(carbon, dcarbon)) +
  geom_point() +
  geom_hline(yintercept = 0, col="red")+
  labs(y="Derivative\n (Rate of change of forest carbon) (kg/C/year)", 
       x="Current Forest Carbon Stock (kgC)")
```

```{r}
# try it out with different initial conditions to watch how the system moves to a stable state
tm = seq(from=1, to=500)
Pinitial = 500
gps = list(harv=10, K=1000, r=0.05, mincarbon=0)

res = ode(Pinitial,tm, fixed_harvest, gps)
colnames(res)=c("time","carbon")

ggplot(as.data.frame(res), aes(time, carbon))+geom_point()+
   labs(y="Forest Biomass (kgC)", x="Year", title="Fixed Harvest of 10kg/year\n Starting conditions 500kgC")
```

```{r}

tm = seq(from=1, to=500)
Pinitial = 15
gps = list(harv=10, K=1000, r=0.05, mincarbon=0)

# try a different method
res = ode(Pinitial,tm, fixed_harvest, gps, method="euler" )
```

```{r}

colnames(res)=c("time","carbon")

ggplot(as.data.frame(res), aes(time, carbon)) +
  geom_point() +
   labs(y="Forest Biomass (kgC)", 
        x="Year", 
        title="Fixed Harvest of 10kg/year\n Starting conditions 1kgC")
```






