---
title: "08_lorenze_equations_chaos"
author: "Steven Cognac"
date: '2022-05-17'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(deSolve)
library(patchwork)

source(here("R/lorenz.R"))
```

Lorenz Equations (for fluid dynamics), P, R, B are parameters (fixed values), x,y,z variables that change with time that describe how conveciton in the atmosphere works - a cell that is warmed from below and cooled from above

Developed by Meteorologist Edward Lorenz - early climate model development in 1960s

 - fluid convection
 - x rate of convective overturning
 - y departure from linear horizontal temperature gradient
 - z departure from linear vertical temperature difference
 
Lorenz equations are example of dynamic systems that can exhibit stable and chaotic states depending on parameters and initial conditions

Lorenze equations useful when you want to model tiny changes in a system and how they can lead to major shifts in systems.

```{r}
# lorenze
pars = list(a=10,b=28,c=8/3)
res = ode(func=lorenz, c(x=0.1,y=0,z=0), times=seq(0,50,by=0.01), parms=pars)

x_change <- ggplot(as.data.frame(res), aes(x,y, col=time)) +
  geom_point() +
  theme(legend.position = "none")
```


```{r}
y_change <- ggplot(as.data.frame(res), aes(x,z, col=time)) +
  geom_point() +
  theme(legend.position = "none")

```

```{r}
z_change <- ggplot(as.data.frame(res), aes(y,z, col=time)) +
  geom_point()
```

```{r}
x_change + y_change + z_change

```


```{r}
ressimple = as.data.frame(res) %>% gather(key="var", value="value",-time)

ggplot(ressimple, aes(time, value, col=var)) +
  geom_line()
```

